---
title: "Signalling pathways in PCAWG"
output:
  html_document:
    df_print: paged
---

```{r}
knitr::opts_chunk$set(comment = '', fig.width = 6, 
                      fig.height = 6, echo = FALSE, 
                      warning = FALSE)
```


```{r load-packages, message=FALSE}
library(here)
fig.dir = here("figures/pathways_analysis")

if(! file.exists(fig.dir)){
    dir.create(fig.dir, recursive = TRUE)
}

source(here("R/load_packages.R"))
```

```{r loading}
mutated.pathways.tissues = readRDS(file = here("data/RDS/PCAWG/10_onco_pathways",
                                              "pcawg_pathways.RDS"))

PCAWG.full.subset.ann.pathways = readRDS(file = here("data/RDS/PCAWG/10_onco_pathways",
                                              "PCAWG.full.subset.ann.pathways.RDS"))
```

### Heatmap of cancer types and pathways

```{r pathway-summary-heatmap}
pathway.summaries.by.tissues = mutated.pathways.tissues %>% 
    mutate(across(`Cell Cycle`:Telomeric, ~as.numeric(.x > 0))) %>% 
    group_by(Cancer.Types) %>% 
    summarize(across(`Cell Cycle`:Telomeric, sum))

pcawg.pathway.summaries.gg = pathway.summaries.by.tissues  %>% 
    gather("Pathway", "Activity", -Cancer.Types) %>% 
    group_by(Cancer.Types, Pathway) %>% 
    summarise(total = sum(Activity))

path.order = pcawg.pathway.summaries.gg %>% 
    dplyr::select(Pathway, total) %>% 
    group_by(Pathway) %>% 
    summarise(ss = sum(total)) %>% 
    arrange(desc(ss)) %>% 
    pull(Pathway)

pcawg.pathways.plot = pcawg.pathway.summaries.gg %>% rename(Counts = total) %>% 
    mutate(Pathway = factor(Pathway, levels = path.order)) %>% 
    
    ggplot(aes(x = Pathway, y = Cancer.Types) ) + 
        geom_tile(aes(fill = Counts), color = "gray80") +
        geom_text(aes(label = Counts), size = 3.4) + 
    #scale_fill_viridis(option = "plasma")
    scale_fill_gradient(low = "aliceblue", high = "orange") + 
    theme_minimal(base_size = 14) + 
    xlab("") + ylab("") + 
    ggtitle("Pathway alterations across cancer types",) +
    theme(panel.grid = element_blank(),
          axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
          legend.position = "none",
          plot.title = element_text(size = 12))
   

print(pcawg.pathways.plot)

ggsave(plot = pcawg.pathways.plot, 
       filename = file.path(fig.dir, "pcawg_pathways_summary.png"),
      width = 5, height = 6.2) 

ggsave(plot = pcawg.pathways.plot, 
       filename = file.path(fig.dir, "pcawg_pathways_summary.pdf"),
      width = 5, height = 6.2) 

```

### Interactions between PCAWG signatures and 10 pathway alterations

```{r path-sig-prep}
PATH_MIN_TISSUES = 30

abundant.tissues = which(PCAWG.full.subset.ann.pathways$Cancer.Types %>% 
                             table() > PATH_MIN_TISSUES) %>% names()
```

#### Interactions between all pathways and signatures

```{r}
all.path.sig.assocs = get_sig_path_assocs(
    PCAWG.full.subset.ann.pathways %>% dplyr::select(4:ncol(.)),
    mutated.pathways.tissues %>% dplyr::select(4:ncol(.)),
    p.val.threshold = 0.05)

all.path.sig.heatmap = myggheatmap(
    dataMatrix = rm_zeros(all.path.sig.assocs), 
    points = T, revColors = F, revRow = T, 
    scaleName = "log(OR)",
    colorPalette = "RdBu",
    limits = c(-5, 5))

all.path.sig.heatmap

ggsave(plot = all.path.sig.heatmap,
       file = file.path(fig.dir, "all.path.sig.heatmap.png"),
       width = 5, height = 6)
```

### Creating a signature-pathway matrix for GES in octave

```{r}
tissue = "Skin_Melanoma"

tissue.sigs.mat = subset_tissue(PCAWG.full.subset.ann.pathways, tissue) %>% 
    select(4:ncol(.))

tissue.sigs.mat = tissue.sigs.mat[,colSums(tissue.sigs.mat > 0) > 10]

tissue.path.mat = subset_tissue(mutated.pathways.tissues, tissue = tissue) %>% 
    select(4:ncol(.))

tissue.path.mat[ tissue.path.mat != 0 ] = 1
tissue.path.mat = tissue.path.mat[, colSums(tissue.path.mat > 0) > 6]

tissue.octave.out = cbind(tissue.sigs.mat, tissue.path.mat)

write_tsv(tissue.octave.out, file = here("data/test/GES_test_melanoma_sig_pathways.tsv"))

```

Reading octave output for melanoma.

```{r}
tissue.octave.adjacency = read.table(here("data/test/GES_melanoma_adjacency.tsv"),
                                     sep = "\t", col.names = colnames(tissue.octave.out),
                                     row.names = colnames(tissue.octave.out))

G = igraph::graph_from_adjacency_matrix(as.matrix(tissue.octave.adjacency))


tbl.G = as_tbl_graph(G, directed = TRUE) %>% 
    activate(nodes) %>% 
    mutate(type = ifelse(name %in% colnames(tissue.sigs.mat), 
                         yes = "signature", no = "pathways")) %>% 
    mutate(node_isolated = node_is_isolated()) %>% 
    filter(!node_isolated)



raw.network = tbl.G  %>% ggraph(layout = "kk") +
    geom_edge_link(arrow = arrow(length = unit(2, units = "mm"), angle = 25),
                   start_cap = circle(3, 'mm'),
                   end_cap = circle(3, 'mm'),
                   color = "grey60") +
    geom_node_point(aes(shape = type), color = "darkblue", size = 3) +
    geom_node_text(aes(label = name), repel = TRUE, force = 60) + 
    theme_void() + ggtitle("Raw signatures")

tissue.logged.octave.adjacency = read.table(here("data/test/GES_melanoma_logged_adjacency.tsv"),
                                     sep = "\t", col.names = colnames(tissue.octave.out),
                                     row.names = colnames(tissue.octave.out))

tbl.logged.G = as_tbl_graph(igraph::graph_from_adjacency_matrix(as.matrix(tissue.octave.adjacency)), directed = TRUE)


logged.network = tbl.logged.G  %>% ggraph(layout = "kk") +
    geom_edge_link(arrow = arrow(length = unit(2, units = "mm"), angle = 25),
                   end_cap = circle(3, 'mm'),
                   color = "grey60") +
    geom_node_point(color = "red") +
    geom_node_text(aes(label = name), repel = TRUE) + 
    theme_void() + ggtitle("Logged signatures")

comp.network =ggarrange(logged.network, raw.network)

```



#### Fisher's exact test between signatures and pathways in individual tissues

```{r, warning = FALSE}
tissue.odds.ratio = pcawg_sig_pathway_int(sigs.input = PCAWG.full.subset.ann.pathways,
                                 pathways.input = mutated.pathways.tissues,
                                 interaction_function = get_sig_path_assocs, 
                                 path.min.tissues = 30,
                                 p.val.threshold = 0.1,
                                 p.adjust = TRUE,
                                 method = "BH")

for (tissue in names(tissue.odds.ratio)) {
   
    if (any(abs(tissue.odds.ratio[[tissue]]) > 0) ) {
        cat(tissue, "\n")
        int.heatmap = ggheatmap_wrapper(tissue.odds.ratio[[tissue]], 
                                        title = tissue)
        
        ggsave(plot = int.heatmap,
               filename = file.path(fig.dir, "pathway_to_signature", "by_tissues", "oddsratio",  
                                    paste0(tissue, "_sig_pathway.dotplot.png")),
               height = 5, width = 6, bg = "white")
    }
}
```

### Linear regression predicting signatures from pathways across tissues

I tried logging the signatures before the linear regression or using raw counts.
In this first case, I tried logging. 


```{r, warning = FALSE}
tissue.lm.coefs.mats.log = pcawg_sig_pathway_int(sigs.input = PCAWG.full.subset.ann.pathways,
                                 pathways.input = mutated.pathways.tissues,
                                 interaction_function = get_sig_path_lms, 
                                 robust = FALSE,
                                 path.min.tissues = 30,
                                 p.val.threshold = 0.1,
                                 p.adjust = TRUE,
                                 method = "BH",
                                 sig.log = TRUE,
                                 path.to.sig = TRUE)
 
for (tissue in abundant.tissues) {
    cat(tissue, "\n")   
    if (any(abs(tissue.lm.coefs.mats.log[[tissue]]) > 0) ) {
        int.heatmap = ggheatmap_wrapper(tissue.lm.coefs.mats.log[[tissue]], 
                                        title = tissue)
        
        ggsave(plot = int.heatmap,
               filename = file.path(fig.dir, "pathway_to_signature",
                                    "by_tissues", "lm_log", 
                                    paste0(tissue, "_sig_tilde_pathway.dotplot.png")),
               height = 5, width = 6, bg = "white")
    }
}
```

In the example below I'm trying with raw counts.

```{r, warning = FALSE}
tissue.lm.coefs.mats = pcawg_sig_pathway_int(sigs.input = PCAWG.full.subset.ann.pathways,
                                 pathways.input = mutated.pathways.tissues,
                                 interaction_function = get_sig_path_lms, 
                                 robust = FALSE,
                                 path.min.tissues = 30,
                                 p.val.threshold = 0.05,
                                 p.adjust = TRUE,
                                 method = "BH",
                                 sig.log = FALSE)
    

for (tissue in abundant.tissues) {
   
    if (any(abs(tissue.lm.coefs.mats[[tissue]]) > 0) ) {
        cat(tissue, "\n")
        int.heatmap = ggheatmap_wrapper(tissue.lm.coefs.mats[[tissue]], 
                                        title = tissue)
        
        ggsave(plot = int.heatmap,
               filename = file.path(fig.dir,"pathway_to_signature", "by_tissues", "lm_counts", 
                                    paste0(tissue, "_sig_tilde_pathway.dotplot.png")),
               height = 5, width = 6, bg = "white")
    }
}
```

#### Summarizing the regression of signatures ~ pathways

```{r, fig.width=6, fig.height=4, warning= FALSE}
pp.logged = plot_all_counts(tissue.lm.coefs.mats.log, threshold = 0.2, psize = 5)
print(pp.logged)

ggsave(plot = pp.logged,
               filename = file.path(fig.dir, "pathway_to_signature", 
                                    "by_tissues", "lm_log", 
                                    "logged_summarized_OLS.png"),
               height = 2.3, width = 2.9, bg = "white")

ggsave(plot = pp.logged,
               filename = file.path(fig.dir, "pathway_to_signature", 
                                    "by_tissues", "lm_log", "logged_summarized_OLS_path_to_sigs.pdf"),
               height = 2.3, width = 2.9, bg = "white", device = cairo_pdf, family = "DejaVu Sans")

```


```{r, fig.width=6, fig.height=4, warning= FALSE}
pp.logged.excl.melanoma = plot_all_counts(
    tissue.lm.coefs.mats.log[setdiff(names(tissue.lm.coefs.mats.log), "Skin_Melanoma")],
    threshold = 0.2, psize = 5)
print(pp.logged.excl.melanoma)

ggsave(plot = pp.logged.excl.melanoma,
               filename = file.path(fig.dir, "pathway_to_signature", 
                                    "by_tissues", "lm_log", 
                                    "logged_summarized_OLS_excl_melanoma_path_to_sigs.png"),
               height = 2, width = 2.4, bg = "white")

ggsave(plot = pp.logged.excl.melanoma,
               filename = file.path(fig.dir, "pathway_to_signature", 
                                    "by_tissues", "lm_log", "logged_summarized_OLS_excl_melanoma_path_to_sigs.pdf"),
               height = 6, width = 8, bg = "white", device = cairo_pdf, family = "DejaVu Sans")

```
We can look at some select interactions. E.g. tissues where the interaction
between NOTCH and Aging is observed. 

```{r}
NOTCH.Aging = summarize_int_mat(tissue.lm.coefs.mats.log)["Ageing", "NOTCH"]
cat(gsub(",", ", ", NOTCH.Aging))
```

NOTCH and SBS40

```{r}
NOTCH.SBS40 = summarize_int_mat(tissue.lm.coefs.mats.log)["SBS40", "NOTCH"]
cat(gsub(",", ", ", NOTCH.SBS40))
```

```{r}
WNT.SBS8 = summarize_int_mat(tissue.lm.coefs.mats.log)["SBS8", "WNT"]
cat(gsub(",", ", ", WNT.SBS8))
```

```{r}
TP53.Ageing = summarize_int_mat(tissue.lm.coefs.mats.log)["Ageing", "TP53"]
cat(gsub(",", ", ", TP53.Ageing))
```

```{r}
TP53.ROS = summarize_int_mat(tissue.lm.coefs.mats.log)["ROS", "TP53"]
cat(gsub(",", ", ", TP53.ROS))
```


```{r}
TP53.HR = summarize_int_mat(tissue.lm.coefs.mats.log, pos.ints = FALSE)["HR", "TP53"]
cat(gsub(",", ", ", TP53.HR))
```

### Case inspection of melanoma

```{r}
tissue = "Skin_Melanoma"
skin.data = get_tissue_pathway_activities(tissue, 
                                          sigs.input = PCAWG.full.subset.ann.pathways,
                                          pathways.input = mutated.pathways.tissues)

skin.data$sigs.logged = skin.data$sigs %>% 
    mutate(across(.cols = everything(), ~ log(.x + 1 )))

skin.concat = merge(skin.data$sigs.logged, skin.data$paths, by = "row.names")


uv.rtkras = skin.concat %>% 
    mutate (`RTK RAS` = factor(`RTK RAS`, levels = c(0, 1,2,3))) %>%
    ggplot(aes(x = `RTK RAS`, y = UV, 
                           group = `RTK RAS`, fill = `RTK RAS`) ) + 
    geom_boxplot() + geom_jitter(width = 0.15, height = 0.15) + 
    theme_bw(base_size = 13) + 
    ylab("log(UV)")

print(uv.rtkras)

ggsave(filename = file.path(fig.dir, paste0(tissue, "uv_rtkras_relationship.png")), 
       plot = uv.rtkras, width = 4, height = 3.2)


rev.uv.rtkras = skin.concat %>% 
    mutate (`RTK RAS` = factor(`RTK RAS`, levels = c(0, 1,2,3))) %>%
    ggplot(aes(x = UV, y = `RTK RAS`) ) + 
    geom_jitter(width = 0.1, height = 0.15) + 
    theme_bw(base_size = 13) + 
    xlab("log(UV)")

ggsave(filename = file.path(fig.dir, paste0(tissue, "rev_uv_rtkras_relationship.png")), 
       plot = rev.uv.rtkras, width = 4, height = 3.2)
```


```{r}
skin.hippo.apo = skin.concat %>% 
    ggplot(aes(x = factor(HIPPO), y = APOBEC, 
                            fill = factor(HIPPO) ) ) +
    theme_bw(base_size = 13) +
    geom_boxplot(outlier.shape = NA) + 
    geom_jitter(width = 0.2, height = 0.02) +
    geom_smooth(aes(group = 1), 
                    data = skin.concat, method = lm, se = FALSE)

ggsave(filename = file.path(fig.dir, paste0(tissue, "apobec_hippo_relationship.png")), 
       plot = skin.hippo.apo, width = 4, height = 3.2)
```

```{r}
skin.hippo.apo.rev = skin.concat %>% 
    ggplot(aes(x = APOBEC, y = HIPPO)) +
    theme_classic() + 
    scale_x_continuous(expand = c(0.2, 0.2)) + 
    geom_smooth(method = "glm", 
    method.args = list(family = "binomial"), 
    se = FALSE) + geom_jitter(aes(color = factor(HIPPO)), width = 0.1, height = 0.1)

ggsave(filename = file.path(fig.dir, paste0(tissue, "apobec_hippo_rev_relationship_logistic.png")), 
       plot = skin.hippo.apo.rev, width = 4, height = 3.2)
```




### Logistic regression predicting pathways from signatures across tissues

```{r}
tissue.log.coefs.mats.log = pcawg_sig_pathway_int(sigs.input = PCAWG.full.subset.ann.pathways,
                                 pathways.input = mutated.pathways.tissues,
                                 interaction_function = get_sig_path_lms, 
                                 path.min.tissues = 30,
                                 p.val.threshold = 0.1,
                                 p.adjust = TRUE,
                                 method = "BH",
                                 sig.log = TRUE,
                                 robust = FALSE,
                                 path.to.sig = FALSE)
 
   
for (tissue in abundant.tissues) {
    if (any(abs(tissue.log.coefs.mats.log[[tissue]]) > 0) ) {
        cat(tissue, "\n")
        int.heatmap = ggheatmap_wrapper(tissue.log.coefs.mats.log[[tissue]], 
                                        title = tissue)
        
        ggsave(plot = int.heatmap,
               filename = file.path(fig.dir, "signature_to_pathway", 
                                    "by_tissues", "log_log", 
                                    paste0(tissue, "_path_tilde_sig.dotplot.png")),
               height = 5, width = 6, bg = "white")
    }
}

```

#### Summarizing the regression of pathways ~ signatures

```{r}
pp.log.log = plot_all_counts(tissue.log.coefs.mats.log, threshold = 0.1, psize = 6,
                             expand.mult = c(0.13, 0.13))
print(pp.log.log)

ggsave(plot = pp.log.log,
               filename = file.path(fig.dir, "signature_to_pathway", 
                                    "by_tissues", "log_log", 
                                    "logged_summarized.png"),
               height = 1.15, width = 1.6, bg = "white")
```

```{r}

```

