---
title: "Survival summaries"
author: "Anna Hakobyan"
date: "2022-11-20"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(comment = '', fig.width = 6, 
                      fig.height = 6, echo = FALSE, 
                      warning = FALSE)
```



```{r load-packages, include = FALSE}
library(here)
library(survival)
library(survminer)

source(here("R/load_packages.R"))
library(gridExtra)
library(ggforestplot)

fig.dir = here("figures/survival_summaries")

```


```{r}
pcawg.sig.sig.pos = readRDS(here("data/RDS/PCAWG/sig_sig_survivals/pos.coxlist.totmut.nonbinary.RDS"))
pcawg.sig.sig.neg = readRDS(here("data/RDS/PCAWG/sig_sig_survivals/neg.coxlist.totmut.nonbinary.RDS"))

tcga.sig.sig.pos = readRDS(here("data/RDS/TCGA/sig_sig_survivals/pos.coxlist.totmut.nonbinary.RDS"))
tcga.sig.sig.neg = readRDS(here("data/RDS/TCGA/sig_sig_survivals/neg.coxlist.totmut.nonbinary.RDS"))

pcawg.sig.path.pos = readRDS(here("data/RDS/PCAWG/sig_pathway_survivals/pos.coxlist.totmut.nonbinary.RDS"))
pcawg.sig.path.neg = readRDS(here("data/RDS/PCAWG/sig_pathway_survivals/neg.coxlist.totmut.nonbinary.RDS"))

tcga.sig.path.pos = readRDS(here("data/RDS/TCGA/sig_pathway_survivals/pos.coxlist.totmut.nonbinary.RDS"))
tcga.sig.path.neg = readRDS(here("data/RDS/TCGA/sig_pathway_survivals/neg.coxlist.totmut.nonbinary.RDS"))


# pcawg.sig.sig.pos = readRDS(here("data/RDS/PCAWG/sig_sig_survivals/pos.coxlist.nototmut.nonbinary.RDS"))
# pcawg.sig.sig.neg = readRDS(here("data/RDS/PCAWG/sig_sig_survivals/neg.coxlist.nototmut.nonbinary.RDS"))
# 
# tcga.sig.sig.pos = readRDS(here("data/RDS/TCGA/sig_sig_survivals/pos.coxlist.nototmut.nonbinary.RDS"))
# tcga.sig.sig.neg = readRDS(here("data/RDS/TCGA/sig_sig_survivals/neg.coxlist.nototmut.nonbinary.RDS"))
# 
# pcawg.sig.path.pos = readRDS(here("data/RDS/PCAWG/sig_pathway_survivals/pos.coxlist.nototmut.nonbinary.RDS"))
# pcawg.sig.path.neg = readRDS(here("data/RDS/PCAWG/sig_pathway_survivals/neg.coxlist.nototmut.nonbinary.RDS"))
# 
# tcga.sig.path.pos = readRDS(here("data/RDS/TCGA/sig_pathway_survivals/pos.coxlist.nototmut.nonbinary.RDS"))
# tcga.sig.path.neg = readRDS(here("data/RDS/TCGA/sig_pathway_survivals/neg.coxlist.nototmut.nonbinary.RDS"))
```

### Forest plot for HR for tissues - Age and TMB


```{r}
pcawg.pos.sig.sig.HRs.all = HR_summary_for_all(pcawg.sig.sig.pos, type = "pos")
pcawg.neg.sig.sig.HRs.all = HR_summary_for_all(pcawg.sig.sig.neg, type = "neg")

pcawg.pos.sig.path.HRs.all = HR_summary_for_all(pcawg.sig.path.pos, type = "pos")
pcawg.neg.sig.path.HRs.all = HR_summary_for_all(pcawg.sig.path.neg, type = "neg")

tcga.pos.sig.sig.HRs.all = HR_summary_for_all(tcga.sig.sig.pos, type = "pos")
tcga.neg.sig.sig.HRs.all = HR_summary_for_all(tcga.sig.sig.neg, type = "neg")

tcga.pos.sig.path.HRs.all = HR_summary_for_all(tcga.sig.path.pos, type = "pos")
tcga.neg.sig.path.HRs.all = HR_summary_for_all(tcga.sig.path.neg, type = "neg")
```


```{r}
HR.all.vars = list(pcawg.pos.sig.sig.HRs.all = pcawg.pos.sig.sig.HRs.all, 
                   pcawg.neg.sig.sig.HRs.all = pcawg.neg.sig.sig.HRs.all,
                     pcawg.pos.sig.path.HRs.all = pcawg.pos.sig.path.HRs.all, 
                     pcawg.neg.sig.path.HRs.all = pcawg.neg.sig.path.HRs.all, 
                     tcga.pos.sig.sig.HRs.all = tcga.pos.sig.sig.HRs.all,
                     tcga.neg.sig.sig.HRs.all = tcga.neg.sig.sig.HRs.all, 
                     tcga.pos.sig.path.HRs.all = tcga.pos.sig.path.HRs.all, 
                     tcga.neg.sig.path.HRs.all = tcga.neg.sig.path.HRs.all)

HR.all.vars.total.plots = list()
# 
# for (var in names(HR.all.vars)) {
#     HR.all.vars.total.plots[[var]] = HR.all.vars[[var]]  %>%
#         filter(P.val < 0.05) %>%
#    ggplot(aes(y = params, x = estimate, color = tissue)) +
#     geom_hline(aes(yintercept = params), alpha = 0.1) +
#       theme_classic(base_size = 15) +
#     geom_point(position=position_dodge(0.7), shape = 15, size = 3) +
#     geom_errorbar(aes(y = params, xmin = lower.95, xmax = upper.95),
#                   width=.2,
#                   position=position_dodge(0.7)) +
#     scale_y_discrete(position = "right") +
#     geom_vline(xintercept = 1, linetype="dashed", color = "gray") +
#     # coord_cartesian(xlim = c(-1, 20)) +
# 
#   theme(axis.line.y = element_blank(),
#         axis.ticks.y = element_blank()) + ggtitle(var)
# }


for (var in names(HR.all.vars)) {
    HR.all.vars.total.plots[[var]] = HR.all.vars[[var]]  %>% drop_na() %>%
        filter(P.val < 0.05) %>%
   ggplot(aes(y = params, x = log(estimate), color = tissue)) +
    geom_hline(aes(yintercept = params), alpha = 0.1) +
      theme_classic(base_size = 15) +
    geom_point(position=position_dodge(0.7), shape = 15, size = 3) +
    geom_errorbar(aes(y = params, xmin = log(lower.95), xmax = log(upper.95)),
                  width=.2,
                  position=position_dodge(0.7)) +
    scale_y_discrete(position = "right") +
    geom_vline(xintercept = 0, linetype="dashed", color = "gray") +
    # coord_cartesian(xlim = c(-1, 20)) +

  theme(axis.line.y = element_blank(),
        axis.ticks.y = element_blank()) + ggtitle(var)
}

pdf(file.path(fig.dir, "logHR.all.survival.summaries.raw.pdf"),
    width = 7, height = 3.5)
for (var in names(HR.all.vars.total.plots)) {
    print (HR.all.vars.total.plots[[var]])
}
dev.off()



for (var in names(HR.all.vars)) {
    HR.all.vars.total.plots[[var]] = HR.all.vars[[var]]  %>% drop_na() %>%
        filter(P.val < 0.05) %>%
        filter(! params %in% c("age_at_diagnosis", "log(total_muts + 1)")) %>%
   ggplot(aes(y = params, x = log(estimate), color = tissue)) +
    geom_hline(aes(yintercept = params), alpha = 0.1) +
      theme_classic(base_size = 15) +
    geom_point(position=position_dodge(0.7), shape = 15, size = 3) +
    geom_errorbar(aes(y = params, xmin = log(lower.95), xmax = log(upper.95)),
                  width=.2,
                  position=position_dodge(0.7)) +
    scale_y_discrete(position = "right") +
    geom_vline(xintercept = 0, linetype="dashed", color = "gray") +
        xlab("log(HR)") +
    # coord_cartesian(xlim = c(-1, 20)) +

  theme(axis.line.y = element_blank(),
        axis.ticks.y = element_blank()) + ggtitle(var)
}

pdf(file.path(fig.dir, "logHR.subset.survival.summaries.raw.pdf"),
    width = 7, height = 3.5)
for (var in names(HR.all.vars.total.plots)) {
    print (HR.all.vars.total.plots[[var]])
}
dev.off()


```


## PCAWG
#### PCAWG sig-sig interactions

```{r}
# pcawg.pos.sig.sig.HRs.all = HR_summary_for_all(pcawg.sig.sig.pos)
# pcawg.neg.sig.sig.HRs.all = HR_summary_for_all(pcawg.sig.sig.neg)

pcawg.sig.sig.comb = rbind (pcawg.pos.sig.sig.HRs.all, pcawg.neg.sig.sig.HRs.all)
pcawg.sig.sig.comb$data = "PCAWG"

p = plot_param_piechart(pcawg.sig.sig.comb, param = "log(total_muts + 1)")

ggsave(plot = p, filename = file.path(fig.dir, "pcawg.sig.sig.tmb.piechart.pdf"),
       width = 3, height = 1.5)
```

```{r}
p = plot_param_piechart(pcawg.sig.sig.comb, param = "age_at_diagnosis")

ggsave(plot = p, filename = file.path(fig.dir, "pcawg.sig.sig.age.piechart.pdf"),
       width = 3, height = 1.5)
```


```{r}
pcawg.age.ggforest = plot_HR_vars(
    all.conds.df = pcawg.sig.sig.comb, param = "age_at_diagnosis",
             average = TRUE, no_stripes = FALSE, log.HR = FALSE) + 
    theme(axis.title.y = element_blank())

ggsave(plot = pcawg.age.ggforest, 
       filename = file.path(fig.dir, "pcawg.age.ggforest.pdf"),
       width = 4.5, height = 3)


pcawg.age.ggforest.log.HR = plot_HR_vars(
    all.conds.df = pcawg.sig.sig.comb, param = "age_at_diagnosis", log.HR = TRUE,
             average = TRUE, no_stripes = FALSE)

ggsave(plot = pcawg.age.ggforest.log.HR, 
       filename = file.path(fig.dir, "pcawg.age.ggforest.log.HR.pdf"),
       width = 4.5, height = 3)

pcawg.tmb.ggforest = plot_HR_vars(
    all.conds.df = pcawg.sig.sig.comb, param = "log(total_muts + 1)",
             average = TRUE, no_stripes = FALSE)

ggsave(plot = pcawg.tmb.ggforest, 
       filename = file.path(fig.dir, "pcawg.tmb.ggforest.pdf"),
       width = 4.5, height = 3)

pcawg.tmb.ggforest.log.HR = plot_HR_vars(
    all.conds.df = pcawg.sig.sig.comb, param = "log(total_muts + 1)", log.HR = TRUE,
             average = TRUE, no_stripes = FALSE)

ggsave(plot = pcawg.tmb.ggforest.log.HR, 
       filename = file.path(fig.dir, "pcawg.tmb.ggforest.log.HR.pdf"),
       width = 4.5, height = 3)
```



```{r}
pp = plot_sigint_forest(pcawg.sig.sig.comb)

ggsave(plot = pp, filename = file.path(fig.dir, "pcawg.sig.sig.significant_interaction.pdf"),
       width = 7, height = 2.8)
```
 
#### PCAWG sig-path interactions

```{r}
# pcawg.pos.sig.sig.HRs.all = HR_summary_for_all(pcawg.sig.sig.pos)
# pcawg.neg.sig.sig.HRs.all = HR_summary_for_all(pcawg.sig.sig.neg)

pcawg.sig.path.comb = rbind (pcawg.pos.sig.path.HRs.all, pcawg.neg.sig.path.HRs.all)


pcawg.sig.path.age.tissues = pcawg.sig.path.comb %>% 
    filter(params == "age_at_diagnosis") %>% 
    mutate(HR.sign = estimate > 1,
           significance = ifelse(sig.star != " ", TRUE, FALSE),
           HR.sign = ifelse(significance == FALSE, FALSE, HR.sign)) %>% 
    group_by(tissue, significance, HR.sign) %>% 
    select(tissue, significance, HR.sign) %>% unique()

p = pcawg.sig.path.age.tissues %>% 
    group_by(significance, HR.sign) %>% 
    summarize(counts = n()) %>%
    ggplot(aes(x = "", y = counts, fill = significance)) +
    geom_col(color = "white") +
    geom_text(aes(label = counts),
              position = position_stack(vjust = 0.5), size =15) +
    scale_fill_manual(values = c("gray90", "coral1")) +
    coord_polar(theta = "y") + theme_void() + 
    theme(legend.position = "none")

ggsave(plot = p, filename = file.path(fig.dir, "pcawg.sig.path.age.piechart.pdf"),
       width = 3, height = 3)
```


<!-- ```{r} -->
<!-- pp = plot_sigint_forest(pcawg.sig.path.comb) -->

<!-- ggsave(plot = pp, filename = file.path(fig.dir, "pcawg.sig.path.significant_interaction.pdf"), -->
<!--        width = 7, height = 2.3) -->
<!-- ``` -->
 
 
## TCGA
### TCGA sig-sig interactions


```{r}
tcga.sig.sig.comb = rbind (tcga.pos.sig.sig.HRs.all, tcga.neg.sig.sig.HRs.all)
tcga.sig.sig.comb$data = "TCGA"
 
p = plot_param_piechart(tcga.sig.sig.comb, param = "log(total_muts + 1)")

ggsave(plot = p, filename = file.path(fig.dir, "tcga.sig.sig.tmb.piechart.pdf"),
       width = 3, height = 1.5)
```

```{r}
p = plot_param_piechart(tcga.sig.sig.comb, param = "age_at_diagnosis")

ggsave(plot = p, filename = file.path(fig.dir, "tcga.sig.sig.age.piechart.pdf"),
       width = 3, height = 1.5)
```



```{r}
tcga.age.ggforest = plot_HR_vars(all.conds.df = tcga.sig.sig.comb, param = "age_at_diagnosis",
             average = TRUE, no_stripes = FALSE) + 
    theme(axis.title.y = element_blank())

ggsave(plot = tcga.age.ggforest, 
       filename = file.path(fig.dir, "tcga.age.ggforest.pdf"),
       width = 4.5, height = 3)


tcga.age.ggforest.log.HR = plot_HR_vars(
    all.conds.df = tcga.sig.sig.comb, param = "age_at_diagnosis", log.HR = TRUE,
             average = TRUE, no_stripes = FALSE)

ggsave(plot = tcga.age.ggforest.log.HR, 
       filename = file.path(fig.dir, "tcga.age.ggforest.log.HR.pdf"),
       width = 4.5, height = 3)

tcga.tmb.ggforest = plot_HR_vars(all.conds.df = tcga.sig.sig.comb, param = "log(total_muts + 1)",
             average = TRUE, no_stripes = FALSE)

ggsave(plot = tcga.tmb.ggforest, 
       filename = file.path(fig.dir, "tcga.tmb.ggforest.pdf"),
       width = 4.5, height = 3)

tcga.tmb.ggforest.log.HR = plot_HR_vars(
    all.conds.df = tcga.sig.sig.comb, param = "log(total_muts + 1)", log.HR = TRUE,
             average = TRUE, no_stripes = FALSE)

ggsave(plot = tcga.tmb.ggforest.log.HR, 
       filename = file.path(fig.dir, "tcga.tmb.ggforest.log.HR.pdf"),
       width = 4.5, height = 3)
```



```{r}
pp = plot_sigint_forest(tcga.sig.sig.comb)

ggsave(plot = pp, filename = file.path(fig.dir, "tcga.sig.sig.significant_interaction.pdf"),
       width = 7, height = 3)
```
 
#### TCGA sig-path interactions

```{r}
# tcga.pos.sig.sig.HRs.all = HR_summary_for_all(tcga.sig.sig.pos)
# tcga.neg.sig.sig.HRs.all = HR_summary_for_all(tcga.sig.sig.neg)

tcga.sig.path.comb = rbind (tcga.pos.sig.path.HRs.all, tcga.neg.sig.path.HRs.all)

p = plot_param_piechart(tcga.sig.path.comb, "age_at_diagnosis")

ggsave(plot = p, filename = file.path(fig.dir, "tcga.sig.path.age.piechart.pdf"),
       width = 3, height = 3)
```


```{r}
pp = plot_sigint_forest(tcga.sig.path.comb)

ggsave(plot = pp, filename = file.path(fig.dir, "tcga.sig.path.significant_interaction.pdf"),
       width = 7, height = 3)
```
 
 
### Age and TMB for both datasets

```{r}
both.data.comb = rbind(pcawg.sig.sig.comb, tcga.sig.sig.comb)

p = plot_param_piechart(both.data.comb, param = "age_at_diagnosis")

ggsave(plot = p, filename = file.path(fig.dir, "both.sig.sig.age.piechart.pdf"),
       width = 3, height = 1.5)

both.age.ggforest.log.HR = plot_HR_vars(
    all.conds.df = both.data.comb, param = "age_at_diagnosis", log.HR = TRUE,
             average = TRUE, no_stripes = FALSE)

ggsave(plot = both.age.ggforest.log.HR, 
       filename = file.path(fig.dir, "both.age.ggforest.log.HR.pdf"),
       width = 4, height = 5)


p = plot_param_piechart(both.data.comb, param = "log(total_muts + 1)")

ggsave(plot = p, filename = file.path(fig.dir, "both.sig.sig.tmb.piechart.pdf"),
       width = 3, height = 1.5)


both.tmb.ggforest.log.HR = plot_HR_vars(
    all.conds.df = both.data.comb, param = "log(total_muts + 1)", log.HR = TRUE,
             average = TRUE, no_stripes = FALSE)

ggsave(plot = both.tmb.ggforest.log.HR, 
       filename = file.path(fig.dir, "both.tmb.ggforest.log.HR.pdf"),
       width = 4, height = 5)
```
 
### PCAWG all models with significant interaction between sig-sig 

```{r}
pcawg.pos.ss.picked.surv.out = readRDS(file = file.path(here("data/RDS/PCAWG/sig_sig_survivals"), "pcawg.pos.sig.sig.picked.models.RDS"))
# pcawg.pos.ss.picked.models = lapply(pcawg.pos.ss.picked.surv.out, function(x) {x[["model"]]})

pcawg.neg.ss.picked.surv.out = readRDS(file = file.path(here("data/RDS/PCAWG/sig_sig_survivals"), "pcawg.neg.sig.sig.picked.models.RDS"))
# pcawg.neg.ss.picked.models = lapply(pcawg.neg.ss.picked.surv.out, function(x) {x[["model"]]})
```

```{r}
pcawg.pos.ss.picked.summary = HR_summary_for_all(pcawg.pos.ss.picked.surv.out, type = "pos")
pcawg.neg.ss.picked.summary = HR_summary_for_all(pcawg.neg.ss.picked.surv.out, type = "neg")

pcawg.ss.picked.combined = rbind(pcawg.pos.ss.picked.summary, pcawg.neg.ss.picked.summary)

pp = plot_sigint_forest(pcawg.ss.picked.combined)

ggsave(plot = pp, filename = file.path(fig.dir, "pcawg.sig.sig.adjusted.models.significant_interaction.pdf"),
       width = 7, height = 3)
```


### PCAWG all models with significant interaction between sig-path


```{r}
pcawg.pos.sp.picked.surv.out = readRDS(file = file.path(here("data/RDS/PCAWG/sig_pathway_survivals"), "pcawg.pos.sig.path.picked.models.RDS"))
# pcawg.pos.ss.picked.models = lapply(pcawg.pos.ss.picked.surv.out, function(x) {x[["model"]]})

pcawg.neg.sp.picked.surv.out = readRDS(file = file.path(here("data/RDS/PCAWG/sig_pathway_survivals"), "pcawg.neg.sig.path.picked.models.RDS"))
# pcawg.neg.ss.picked.models = lapply(pcawg.neg.ss.picked.surv.out, function(x) {x[["model"]]})
```


```{r}
pcawg.pos.sp.picked.summary = HR_summary_for_all(pcawg.pos.sp.picked.surv.out, type = "pos")

pcawg.neg.sp.picked.summary = HR_summary_for_all(pcawg.neg.sp.picked.surv.out, type = "neg")
# This list is empty

pcawg.sp.picked.combined = rbind(pcawg.pos.sp.picked.summary, pcawg.neg.sp.picked.summary)
# pcawg.sp.picked.combined = pcawg.pos.sp.picked.summary

pp = plot_sigint_forest(pcawg.sp.picked.combined) + 
    theme(plot.margin = margin(1, 0,0,0, "cm"))

ggsave(plot = pp, filename = file.path(fig.dir, "pcawg.sig.path.adjusted.models.significant_interaction.pdf"),
       width = 7, height = 2)
```


### TCGA all models with significant interaction between sig-sig

```{r}
tcga.pos.ss.picked.surv.out = readRDS(file = file.path(here("data/RDS/TCGA/sig_sig_survivals"), "tcga.pos.sig.sig.picked.models.RDS"))
# tcga.pos.ss.picked.models = lapply(tcga.pos.ss.picked.surv.out, function(x) {x[["model"]]})

tcga.neg.ss.picked.surv.out = readRDS(file = file.path(here("data/RDS/TCGA/sig_sig_survivals"), "tcga.neg.sig.sig.picked.models.RDS"))
# tcga.neg.ss.picked.models = lapply(tcga.neg.ss.picked.surv.out, function(x) {x[["model"]]})
```

```{r}
tcga.pos.ss.picked.summary = HR_summary_for_all(tcga.pos.ss.picked.surv.out, type = "pos")
tcga.neg.ss.picked.summary = HR_summary_for_all(tcga.neg.ss.picked.surv.out, type = "neg")

tcga.ss.picked.combined = rbind(tcga.pos.ss.picked.summary, tcga.neg.ss.picked.summary)

clock.like = c("Ageing", "SBS40", "SBS5")
pairs = combn(clock.like, 2)
pair.strings <- c(paste(pairs[1, ], pairs[2, ], sep = " + "), 
                  paste(pairs[2, ], pairs[1, ], sep = " + "))

tcga.ss.picked.combined = tcga.ss.picked.combined %>% filter(! int %in% pair.strings )

pp = plot_sigint_forest(tcga.ss.picked.combined)

ggsave(plot = pp, filename = file.path(fig.dir, "tcga.sig.sig.adjusted.models.significant_interaction.pdf"),
       width = 7, height = 3.2)
```


### TCGA all models with significant interaction between sig-path

```{r}
tcga.pos.sp.picked.surv.out = readRDS(file = file.path(here("data/RDS/TCGA/sig_pathway_survivals"), "tcga.pos.sig.path.picked.models.RDS"))
# tcga.pos.ss.picked.models = lapply(tcga.pos.ss.picked.surv.out, function(x) {x[["model"]]})

tcga.neg.sp.picked.surv.out = readRDS(file = file.path(here("data/RDS/TCGA/sig_pathway_survivals"), "tcga.neg.sig.path.picked.models.RDS"))
# tcga.neg.ss.picked.models = lapply(tcga.neg.ss.picked.surv.out, function(x) {x[["model"]]})
```

 
```{r}
tcga.pos.sp.picked.summary = HR_summary_for_all(tcga.pos.sp.picked.surv.out, type = "pos")

tcga.neg.sp.picked.summary = HR_summary_for_all(tcga.neg.sp.picked.surv.out, type = "neg")

tcga.sp.picked.combined = rbind(tcga.pos.sp.picked.summary, tcga.neg.sp.picked.summary)

pp = plot_sigint_forest(tcga.sp.picked.combined)

ggsave(plot = pp, filename = file.path(fig.dir, "tcga.sig.path.adjusted.models.significant_interaction.pdf"),
       width = 7, height = 3.2)
```

