---
title: "Signalling pathways in TCGA"
output:
  html_document:
    df_print: paged
---

```{r}
rm(list = ls())
.rs.restartR()

knitr::opts_chunk$set(comment = '', fig.width = 6, 
                      fig.height = 6, echo = FALSE, 
                      warning = FALSE)
```


```{r load-packages, message=FALSE}
library(here)
library(cowplot)
fig.dir = here("figures/pathways_analysis")

if(! file.exists(fig.dir)){
    dir.create(fig.dir, recursive = TRUE)
}

out.dir = here("data", "RDS", "TCGA", "10_onco_pathways")

source(here("R/load_packages.R"))
```

```{r loading}
TCGA.pathways.tissues = readRDS(file = here("data/RDS/TCGA/10_onco_pathways",
                                              "tcga_pathways.RDS"))

TCGA.full.subset.ann.pathways = readRDS(file = here("data/RDS/TCGA/10_onco_pathways",
                                              "TCGA.full.subset.ann.pathways.RDS"))

TCGA.full.subset.ann.pathways$Sample.Names = rownames(TCGA.full.subset.ann.pathways)

TCGA.full.subset.ann = readRDS(here("data", "RDS", "TCGA", "signatures",
                                     "TCGA.full.subset.ann.RDS"))

TCGA.full.subset = readRDS(here("data", "RDS", "TCGA", "signatures",
                                     "TCGA.full.subset.RDS"))

TCGA.active.sig.counts = data.frame(Tissues = TCGA.full.subset.ann$Cancer.Types,
                                      active.sig.counts = rowSums(TCGA.full.subset.ann[4:ncol(TCGA.full.subset.ann)] > 0))

TCGA.sig.path.nulls = readRDS(file = here("data/RDS/TCGA/10_onco_pathways",
                                              "TCGA.sig.path.null.dists.RDS"))
```


### TCGA summary statistics about available data

In total we have `r nrow(TCGA.pathways.tissues)` samples and 
`r unique(TCGA.pathways.tissues$Cancer.Types) %>% length()` tissues.

```{r}
# [1] 1552
TCGA.mutated.pathways.occ.plot = TCGA.pathways.tissues %>% 
    mutate(across (`Cell Cycle`:WNT, ~ . > 0)) %>% 
    rowwise() %>% 
    mutate(sum.mut.paths = sum(c_across(`Cell Cycle`:WNT))) %>% 
    ggplot(aes(x = sum.mut.paths, fill = Cancer.Types) ) + 
    geom_histogram(binwidth = 1, color = "gray90") +
    facet_wrap(~Cancer.Types) +
        theme_classic(base_size = 13) + 
        scale_x_continuous(breaks = seq(1, 7)) + 
        coord_cartesian(expand = FALSE) + 
        scale_fill_manual(values = scales::muted(
            rainbow(TCGA.active.sig.counts %>% pull(Tissues) %>% 
                        n_distinct()), l = 50, c = 90 ), guide = "none" ) +
        xlab("Mutated pathways") + ylab("Number of samples") + 
    theme(strip.text.x = element_text(size = 6))


ggsave(file = file.path(fig.dir, "TCGA.mutated.pathways.occurrence.pdf"),
       plot = TCGA.mutated.pathways.occ.plot, 
       width = 5, height = 5)
```


### Heatmap of cancer types and pathways

```{r pathway-summary-heatmap}
pathway.summaries.by.tissues = TCGA.pathways.tissues %>% 
    mutate(across(`Cell Cycle`:WNT, ~as.numeric(.x > 0))) %>% 
    group_by(Cancer.Types) %>% 
    summarize(across(`Cell Cycle`:WNT, ~ sum(.x, na.rm = TRUE)))

TCGA.pathway.summaries.gg = pathway.summaries.by.tissues  %>% 
    gather("Pathway", "Activity", -Cancer.Types) %>% 
    group_by(Cancer.Types, Pathway) %>% 
    summarise(total = sum(Activity))

path.order = TCGA.pathway.summaries.gg %>% 
    dplyr::select(Pathway, total) %>% 
    group_by(Pathway) %>% 
    summarise(ss = sum(total)) %>% 
    arrange(desc(ss)) %>% 
    pull(Pathway)

TCGA.pathways.plot = TCGA.pathway.summaries.gg %>% rename(Counts = total) %>% 
    mutate(Pathway = factor(Pathway, levels = path.order)) %>% 
    
    ggplot(aes(x = Pathway, y = Cancer.Types) ) + 
        geom_tile(aes(fill = Counts), color = "gray80") +
        geom_text(aes(label = Counts), size = 3.4) + 
    #scale_fill_viridis(option = "plasma")
    scale_fill_gradient(low = "aliceblue", high = "orange") + 
    theme_minimal(base_size = 14) + 
    xlab("") + ylab("") + 
    ggtitle("Pathway alterations across cancer types: TCGA",) +
    theme(panel.grid = element_blank(),
          axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
          legend.position = "none",
          plot.title = element_text(size = 12))
   

print(TCGA.pathways.plot)

ggsave(plot = TCGA.pathways.plot, 
       filename = file.path(fig.dir, "TCGA_pathways_summary.png"),
      width = 5, height = 6.2) 

ggsave(plot = TCGA.pathways.plot, 
       filename = file.path(fig.dir, "TCGA_pathways_summary.pdf"),
      width = 5, height = 6.2) 

```

### Interactions between TCGA signatures and 10 pathway alterations

```{r path-sig-prep}
PATH_MIN_TISSUES = 30

abundant.tissues = which(TCGA.full.subset.ann.pathways$Cancer.Types %>% 
                             table() > PATH_MIN_TISSUES) %>% names()

# number of tissues
# length(abundant.tissues)

# number of total samples
TCGA.full.subset.ann.pathways$Cancer.Types %in% abundant.tissues %>% sum()

```


#### Interactions between all pathways and signatures

```{r}
all.path.sig.assocs = get_sig_path_assocs(
    TCGA.full.subset.ann.pathways %>% dplyr::select(4:ncol(.)),
    TCGA.pathways.tissues %>% dplyr::select(4:ncol(.)),
    p.val.threshold = 0.05)

all.path.sig.heatmap = myggheatmap(
    dataMatrix = rm_zeros(all.path.sig.assocs), 
    points = T, revColors = F, revRow = T, 
    scaleName = "log(OR)",
    colorPalette = "RdBu",
    limits = c(-5, 5))

all.path.sig.heatmap

ggsave(plot = all.path.sig.heatmap,
       file = file.path(fig.dir, "TCGA.all.path.sig.heatmap.png"),
       width = 5, height = 6)
```

<!-- ### Creating a signature-pathway matrix for GES in octave -->

<!-- ```{r} -->
<!-- tissue = "Skin_Melanoma" -->

<!-- tissue.sigs.mat = subset_tissue(TCGA.full.subset.ann.pathways, tissue) %>%  -->
<!--     select(4:ncol(.)) -->

<!-- tissue.sigs.mat = tissue.sigs.mat[,colSums(tissue.sigs.mat > 0) > 10] -->

<!-- tissue.path.mat = subset_tissue(TCGA.pathways.tissues, tissue = tissue) %>%  -->
<!--     select(4:ncol(.)) -->

<!-- tissue.path.mat[ tissue.path.mat != 0 ] = 1 -->
<!-- tissue.path.mat = tissue.path.mat[, colSums(tissue.path.mat > 0) > 6] -->

<!-- tissue.octave.out = cbind(tissue.sigs.mat, tissue.path.mat) -->

<!-- write_tsv(tissue.octave.out, file = here("data/test/GES_test_melanoma_sig_pathways.tsv")) -->

<!-- ``` -->

<!-- Reading octave output for melanoma. -->

<!-- ```{r} -->
<!-- tissue.octave.adjacency = read.table(here("data/test/GES_melanoma_adjacency.tsv"), -->
<!--                                      sep = "\t", col.names = colnames(tissue.octave.out), -->
<!--                                      row.names = colnames(tissue.octave.out)) -->

<!-- G = igraph::graph_from_adjacency_matrix(as.matrix(tissue.octave.adjacency)) -->


<!-- tbl.G = as_tbl_graph(G, directed = TRUE) %>%  -->
<!--     activate(nodes) %>%  -->
<!--     mutate(type = ifelse(name %in% colnames(tissue.sigs.mat),  -->
<!--                          yes = "signature", no = "pathways")) %>%  -->
<!--     mutate(node_isolated = node_is_isolated()) %>%  -->
<!--     filter(!node_isolated) -->



<!-- raw.network = tbl.G  %>% ggraph(layout = "kk") + -->
<!--     geom_edge_link(arrow = arrow(length = unit(2, units = "mm"), angle = 25), -->
<!--                    start_cap = circle(3, 'mm'), -->
<!--                    end_cap = circle(3, 'mm'), -->
<!--                    color = "grey60") + -->
<!--     geom_node_point(aes(shape = type), color = "darkblue", size = 3) + -->
<!--     geom_node_text(aes(label = name), repel = TRUE, force = 60) +  -->
<!--     theme_void() + ggtitle("Raw signatures") -->

<!-- tissue.logged.octave.adjacency = read.table(here("data/test/GES_melanoma_logged_adjacency.tsv"), -->
<!--                                      sep = "\t", col.names = colnames(tissue.octave.out), -->
<!--                                      row.names = colnames(tissue.octave.out)) -->

<!-- tbl.logged.G = as_tbl_graph(igraph::graph_from_adjacency_matrix(as.matrix(tissue.octave.adjacency)), directed = TRUE) -->


<!-- logged.network = tbl.logged.G  %>% ggraph(layout = "kk") + -->
<!--     geom_edge_link(arrow = arrow(length = unit(2, units = "mm"), angle = 25), -->
<!--                    end_cap = circle(3, 'mm'), -->
<!--                    color = "grey60") + -->
<!--     geom_node_point(color = "red") + -->
<!--     geom_node_text(aes(label = name), repel = TRUE) +  -->
<!--     theme_void() + ggtitle("Logged signatures") -->

<!-- comp.network =ggarrange(logged.network, raw.network) -->

<!-- ``` -->


### Type of signature-pathway interactions in TCGA

```{r}
tissue.odds.ratio.unadjusted = sig_pathway_int(
    sigs.input = TCGA.full.subset.ann.pathways,
    pathways.input = TCGA.pathways.tissues,
    interaction_function = get_sig_path_assocs, 
    path.min.tissues = 30,
    p.val.threshold = 0.05,
    p.adjust = FALSE,
    method = "BH")


tissue.lm.coefs.mats.log = sig_pathway_int(sigs.input = TCGA.full.subset.ann.pathways,
                                 pathways.input = TCGA.pathways.tissues,
                                 interaction_function = get_sig_path_lms,
                                 robust = FALSE,
                                 path.min.tissues = 30,
                                 p.val.threshold = 0.05,
                                 p.adjust = FALSE,
                                 method = "BH",
                                 sig.log = TRUE,
                                 path.to.sig = TRUE)

tissue.log.coefs.mats.log = sig_pathway_int(sigs.input = TCGA.full.subset.ann.pathways,
                                 pathways.input = TCGA.pathways.tissues,
                                 interaction_function = get_sig_path_lms,
                                 path.min.tissues = 30,
                                 p.val.threshold = 0.05,
                                 p.adjust = FALSE,
                                 method = "BH",
                                 sig.log = TRUE,
                                 robust = FALSE,
                                 path.to.sig = FALSE)
# 
# sig.path.int.funcs = list(
#     OR = OR.func.null,
#     STP = STP.null,
#     PTS = PTS.null
# )
TCGA.sig.path.interactions = list()

for (tissue in names(tissue.odds.ratio.unadjusted)) {
    int.list = list(OR = tissue.odds.ratio.unadjusted[[tissue]],
                    STP = tissue.log.coefs.mats.log[[tissue]],
                    PTS = tissue.lm.coefs.mats.log[[tissue]])
    TCGA.sig.path.interactions[[tissue]] = int.list
}



```



### Filter the interactions based on the null distributions

```{r}
sig.threshold = 0.05
TCGA.sig.path.interactions.significant = TCGA.sig.path.interactions

metric.name.mapping = list(OP = "Fisher's exact test",
                           STP = "glm(pathway ~ signature)",
                           PTS = "lm(signature ~ pathway)")

out.file = here("supp_data", "TCGA.sig.pathway.significant.interactions.xlsx")
file.remove(out.file)

for(tissue in names(TCGA.sig.path.interactions)) {
    
    for (metric.name in names(TCGA.sig.path.interactions[[tissue]])) {
        cat(tissue, "\t", metric.name, "\n")
        out =  TCGA.sig.path.interactions[[tissue]][[metric.name]]
        
        if (length(out) == 0 ) {
            next
        }
        metric.values = out %>% rm_zeros()
        if (length(metric.values) == 0) {next} 
        
        p.values = matrix(NA, nrow = nrow(metric.values), 
                          ncol = ncol(metric.values), 
                          dimnames = (dimnames(metric.values)))
        
        non.zero.values = which(metric.values != 0, arr.ind = TRUE)
        
        for(i in 1:nrow(non.zero.values)) {
            sig = rownames(metric.values)[non.zero.values[i, 1]]
            path = colnames(metric.values)[non.zero.values[i, 2]]
            
            p.values[sig, path] = min(percentile(
                TCGA.sig.path.nulls[[tissue]][[metric.name]][gsub(" ", ".", sig), path, ] ,
                                                  metric.values[sig, path]),
                1 - percentile(TCGA.sig.path.nulls[[tissue]][[metric.name]][gsub(" ", ".", sig), path, ] ,
                                                  metric.values[sig, path]))
            
            if(p.values[sig, path] > sig.threshold) {
                out[sig, path] = 0
                p.values[sig, path] = 1
            }
        }
        TCGA.sig.path.interactions.significant[[metric.name]][[tissue]] = out
        
        
        
        if (!file.exists(out.file)) {
            wb <- createWorkbook()
            cat("File doesn't exist.\n")
        } else { 
            wb = loadWorkbook(out.file)
        }
        
        if (!(tissue %in% names(wb))) {
            cat("Sheet doesn't exist.\n")
            addWorksheet(wb, tissue)
            # writeData(wb = wb, 
            #           sheet = tissue,
            #           x = metric.name.mapping[[metric.name]])
            start.row = 2
        } else {
            cat("Sheet exists.\n")
            start.row = nrow(readWorkbook(out.file, 
                                          sheet = tissue, 
                                          colNames = FALSE, 
                                          skipEmptyRows = FALSE))+5
        }
        
        writeData(wb = wb, 
                  sheet = tissue,
                  x = metric.name.mapping[[metric.name]],
                  colNames = TRUE,
                  startRow = start.row - 1,
                  startCol = 1) 
        cat("File exists.\n")
        
        
        cat(metric.name, "\t", start.row, "\n") 
        
        ### Write the metric values
        writeData(wb = wb, 
                  sheet = tissue,
                  x = metric.values,
                  colNames = TRUE,
                  rowNames = TRUE,
                  startRow = start.row)
        
        ### Write the p-values annotation
        writeData(wb = wb, 
                  sheet = tissue,
                  x = "p-values",
                  colNames = TRUE,
                  startRow = start.row - 1,
                  startCol = ncol(metric.values)+4) 
        
        ### Write the p-values matrix
        
        p.values[is.na(p.values)] = 1
        
        writeData(wb = wb, 
                  sheet = tissue,
                  x = p.values,
                  colNames = TRUE,
                  rowNames = TRUE,
                  startRow = start.row,
                  startCol = ncol(metric.values)+4)
        
        saveWorkbook(wb, file = out.file, overwrite = TRUE)

    }
}


saveRDS(TCGA.sig.path.interactions.significant, here("data/RDS/TCGA/10_onco_pathways",
                                              "TCGA.sig.path.interactions.significant.RDS"))
```


```{r}
pp.OR = plot_all_counts(lapply(TCGA.sig.path.interactions.significant, function(x) x$OR), 
                                              threshold = 0.2, psize = 5,
                                              expand.mult = c(0.03, 0.03)) + 
    theme(axis.text = element_text(family = "Arial"))

ggsave(plot = pp.OR,
               filename = file.path(fig.dir, "pathway_to_signature", "by_tissues", "oddsratio",
                                    "TCGA_summarized_sig_pathway_null_adjusted.dotplot.pdf"),
               height = 2.7, width = 3.8, bg = "white", device = cairo_pdf)





pp.STP = plot_all_counts(lapply(TCGA.sig.path.interactions.significant, function(x) x$STP), 
                         threshold = 0.1, psize = 6,
                         expand.mult = c(0.03, 0.03))  + 
    theme(axis.text = element_text(family = "Arial"))                      

ggsave(plot = pp.STP,
       filename = file.path(fig.dir, "signature_to_pathway",
                            "by_tissues", "log_log",
                            "TCGA_logged_logistic_summarized.null.adjusted.pdf"),
       height = 2.8, width = 4, bg = "white", device = cairo_pdf)




pp.PTS = plot_all_counts(lapply(TCGA.sig.path.interactions.significant, function(x) x$PTS), 
                         threshold = 0.2, psize = 5,
                         expand.mult = c(0.03, 0.03)) + 
    theme(axis.text = element_text(family = "Arial"))


ggsave(plot = pp.PTS,
               filename = file.path(fig.dir, "pathway_to_signature",
                                    "by_tissues", "lm_log", "TCGA_logged_summarized_OLS_path_to_sigs.lm.null.adjusted.pdf"),
               height = 2.7, width = 4, bg = "white", device = cairo_pdf)


```




### Bipartite graph - not used 

```{r}
sig.path.oddsratio.bipartite = minor_plot(plot_bipartite2(lapply(TCGA.sig.path.interactions.significant, function(x) x$OR)),
                                       expand.factor = 0.5, expand.factor.y = 0.01)

ggsave(plot = sig.path.oddsratio.bipartite,
       filename = file.path(fig.dir, "pathway_to_signature", "by_tissues", "oddsratio",
                                    "TCGA_summarized_sig_pathway_adjusted.bipartite.pdf"),
       height = 5, width = 6, bg = "white", device = cairo_pdf)
```

### Tissues of Odds ratio interactions

```{r}
tissues.odds.ratio.lists = get_interaction_tissues(
    lapply(TCGA.sig.path.interactions.significant, function(x) x$OR))

write.table(as.data.frame(tissues.odds.ratio.lists$pos.tissues), 
          file = here("supp_data", "TCGA_sig_pathway_positive_interaction_tissue_summaries.tsv"),
          row.names = TRUE, col.names = NA, sep = "\t", quote = FALSE)
write.table(as.data.frame(tissues.odds.ratio.lists$neg.tissues), 
          file = here("supp_data", "TCGA_sig_pathway_negative_interaction_tissue_summaries.tsv"),
          row.names = TRUE, col.names = NA, sep = "\t", quote = FALSE)

```

#### Plotting the heatmaps

```{r}
tissue.odds.ratio.null.adjusted = lapply(TCGA.sig.path.interactions.significant, function(x)
    x$OR)
for (tissue in names(tissue.odds.ratio.null.adjusted)) {
   
    if (length(tissue.odds.ratio.null.adjusted[[tissue]]) == 0) {
        next
    }
    if (any(abs(tissue.odds.ratio.null.adjusted[[tissue]]) > 0) ) {
        cat(tissue, "\n")
        int.heatmap = ggheatmap_wrapper(tissue.odds.ratio.null.adjusted[[tissue]], 
                                        title = tissue)
        
        print(int.heatmap)
        ggsave(plot = int.heatmap,
               filename = file.path(fig.dir, "pathway_to_signature", "by_tissues", "oddsratio",
                                    paste0("TCGA_", tissue, "_sig_pathway_null_adjusted.dotplot.png")),
               height = 5, width = 6, bg = "white")
    }
}

tissue.lm.null.adjusted = lapply(TCGA.sig.path.interactions.significant, function(x)
    x$PTS)
for (tissue in abundant.tissues) {
    cat(tissue, "\n")
    if (any(abs(tissue.lm.null.adjusted[[tissue]]) > 0) ) {
        int.heatmap = ggheatmap_wrapper(tissue.lm.null.adjusted[[tissue]],
                                        title = tissue)

        ggsave(plot = int.heatmap,
               filename = file.path(fig.dir, "pathway_to_signature",
                                    "by_tissues", "lm_log",
                                    paste0("TCGA_", tissue, "_sig_tilde_pathway.dotplot.png")),
               height = 5, width = 6, bg = "white")
    }
}



tissue.glm.null.adjusted = lapply(TCGA.sig.path.interactions.significant, 
                                  function(x) x$STP)
for (tissue in abundant.tissues) {
    if (any(abs(tissue.glm.null.adjusted[[tissue]]) > 0) ) {
        cat(tissue, "\n")
        int.heatmap = ggheatmap_wrapper(tissue.log.coefs.mats.log[[tissue]],
                                        title = tissue)

        ggsave(plot = int.heatmap,
               filename = file.path(fig.dir, "signature_to_pathway",
                                    "by_tissues", "log_log",
                                    paste0("TCGA_", tissue, "_path_tilde_sig.glm.null.adjusted.dotplot.png")),
               height = 5, width = 6, bg = "white")
    }
}

```

### Melanoma pathways

```{r}
tissue = "Skin_Melanoma"
skin.annotated.heatmap = pathways_signatures_heatmap(
        tissue = tissue,
        signatures = TCGA.full.subset.ann.pathways, border_color = NA,
        pathways = TCGA.pathways.tissues, main = paste0(tissue, " - TCGA")) %>% 
    add_pheatmap_legend_title
    

ggsave(plot = skin.annotated.heatmap,
       file = file.path(fig.dir, "TCGA.skin.annotated.heatmap.pdf"),
       width = 4.5, height = 4)
```


```{r}
tissue = "Skin_Melanoma"
tissue.heatmap = subset_tissue(TCGA.full.subset, tissue = tissue) %>% 
            plot_tissue_heatmap(main = tissue)

```


### Breast_cancer pathways

```{r}
tissue = "Breast_cancer"
skin.annotated.heatmap = pathways_signatures_heatmap(
        tissue = tissue,
        signatures = TCGA.full.subset.ann.pathways, border_color = NA,
        pathways = TCGA.pathways.tissues, main = paste0(tissue, " - TCGA")) %>% 
    add_pheatmap_legend_title
    

ggsave(plot = skin.annotated.heatmap,
       file = file.path(fig.dir, paste0("TCGA.", tissue, ".annotated.heatmap.pdf")),
       width = 4.5, height = 4)
```




### Stomach_AdenoCa pathways

```{r}
tissue = "Stomach_AdenoCa"
skin.annotated.heatmap = pathways_signatures_heatmap(
        tissue = tissue,
        signatures = TCGA.full.subset.ann.pathways, border_color = NA,
        pathways = TCGA.pathways.tissues, main = paste0(tissue, " - TCGA")) %>% 
    add_pheatmap_legend_title
    

ggsave(plot = skin.annotated.heatmap,
       file = file.path(fig.dir, paste0("TCGA.", tissue, ".annotated.heatmap.pdf")),
       width = 4.5, height = 4)


int.heatmap = ggheatmap_wrapper(tissue.odds.ratio.null.adjusted[[tissue]], 
                                        title = tissue)
        
print(int.heatmap)
ggsave(plot = int.heatmap,
       filename = file.path(fig.dir, "pathway_to_signature", "by_tissues", "oddsratio",
                     paste0("TCGA_", tissue, "_sig_pathway.dotplot.pdf")),
height = 5, width = 6, bg = "white")
```

### Three summaries together

```{r}
pp.TCGA.combined.summaries = ggarrange(pp.OR + ggtitle("Co-occurrence"), 
          pp.PTS + ggtitle("lm(signature ~ pathway)"), 
          pp.STP + ggtitle("glm(pathway ~ signature)"), 
          ncol = 1,
          heights = c(1, 1, 1))

pp.TCGA.combined.summaries = annotate_figure(
    pp.TCGA.combined.summaries, top = text_grob("TCGA signature-pathway analysis summaries", 
               color = "darkblue", face = "bold", size = 11))

ggsave(plot = pp.TCGA.combined.summaries,
               filename = file.path(fig.dir, "TCGA_combined_summaries_sig_pathway.pdf"),
               height = 10, width = 4, bg = "white", device = cairo_pdf)
```



```{r}
pp.TCGA.combined.summaries = ggdraw() +
  draw_plot(pp.OR + ggtitle("Co-occurrence"), x = 0, y = 0.7, width = 0.8, height = .3) +
  draw_plot(pp.PTS + ggtitle("lm(signature ~ pathway)"), x = 0, y = .35, width = .8, height = .35) +
  draw_plot(pp.STP + ggtitle("glm(pathway ~ signature)"), x = 0, y = 0, width = 0.8, height = 0.35)

ggsave(plot = pp.TCGA.combined.summaries,
               filename = file.path(fig.dir, "TCGA_combined_summaries_sig_pathway.pdf"),
               height = 9, width = 5, bg = "white", device = cairo_pdf)
```

### All signatures and pathway mutation OR interaction heatmaps

```{r}

tissue.odds.ratio.adjusted = lapply(TCGA.sig.path.interactions.significant, 
                                      function(x) x$OR)

plotlist = list()
for (tissue in abundant.tissues) {
    path.heatmap = pathways_signatures_heatmap(
        tissue = tissue,
        signatures = TCGA.full.subset.ann.pathways, border_color = NA,
        pathways = TCGA.pathways.tissues, main = tissue)
    
    if (any(abs(tissue.odds.ratio.adjusted[[tissue]]) > 0) ) {
        cat(tissue, "\n")
        int.heatmap = ggheatmap_wrapper(tissue.odds.ratio.adjusted[[tissue]], 
                                        title = tissue)
    } else {
        int.heatmap = grob()
    }
    combined.out = ggarrange(path.heatmap[[4]], int.heatmap, widths = c(1, 1))
    plotlist[[tissue]] = combined.out
}


pdf(file.path(fig.dir, "TCGA_pathways_signatures_heatmaps.pdf"),
    width = 10, height = 4)
for (tissue in names(plotlist)) {
    print(plotlist[[tissue]])
}
dev.off()
```
